---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-", fig.align='center')

```
# EMtree

[![R build status](https://github.com/Rmomal/EMtree/workflows/R-CMD-check/badge.svg)](https://github.com/Rmomal/EMtree/actions)
[![Codecov test coverage](https://codecov.io/gh/Rmomal/EMtree/branch/master/graph/badge.svg)](https://codecov.io/gh/Rmomal/EMtree?branch=master)
[![DOI](https://zenodo.org/badge/166967948.svg)](https://zenodo.org/badge/latestdoi/166967948)

> `EMtree` infers direct species association networks, implementing the procedure described in [Momal *et al.*](https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/2041-210X.13380). This package uses averages over spanning trees within an Expectation-Maximization algorithm to infer conditional dependence networks, and involves plotting functionalities (using `ggraph` and `tydigraph`). 

By default, it uses the Poisson log-Normal Model ([PLNmodels](https://github.com/jchiquet/PLNmodels>)) to accommodate abundance data. However, EMtree is an inference method which only requires an estimate of a Gaussian covariance matrix, and can be used with any model which either use Gaussian latent variables, Gaussian copulas, or Gaussian data transformations.

## Installation

`EMtree` requires  R>3.5.

### CRAN dependencies

```{r,eval=FALSE}
required_CRAN <- c("Matrix", "parallel",  "mvtnorm", "vegan",
                   "ggplot2", "magrittr", "dplyr", "tibble",
                   "PLNmodels","ggraph", "tidygraph","huge")
not_installed_CRAN <- setdiff(required_CRAN, rownames(installed.packages()))
if (length(not_installed_CRAN) > 0) install.packages(not_installed_CRAN)
```

### Installation of `EMtree`
You can install the development version of EMtree:

``` r
devtools::install_github("Rmomal/EMtree")
```

## Reference
Please cite our work using the following reference:

- Momal, R., Robin, S., & Ambroise, C. (2020). Tree‐based inference of species interaction networks from abundance data. *Methods in Ecology and Evolution*, 11(5), 621-632.

