---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-", fig.align='center')

```
# EMtree

[![Travis build status](https://travis-ci.org/Rmomal/EMtree.svg?branch=master)](https://travis-ci.org/Rmomal/EMtree)
[![Codecov test coverage](https://codecov.io/gh/Rmomal/EMtree/branch/master/graph/badge.svg)](https://codecov.io/gh/Rmomal/EMtree?branch=master)
[![DOI](https://zenodo.org/badge/166967948.svg)](https://zenodo.org/badge/latestdoi/166967948)

> EMtree infers interaction networks from abundance data. It uses averages over spanning trees within a Poisson log-Normal Model ([PLNmodels](https://github.com/jchiquet/PLNmodels>)), and involves plotting funcitonalities (using `ggraph` and `tydigraph`).

## Installation

You can install the development version of EMtree:

``` r
devtools::install_github("Rmomal/EMtree")
```

## Example with Fatala river fishes

This is a basic example which shows you how to infer a network, using Barans95 data from the `ade4` package.

### Data


```{r fish_data , message=FALSE}
library(ade4)
library(tidyverse)
data(baran95)
counts = as.matrix(baran95$fau)
covar = as_tibble(baran95$plan)

n = nrow(counts)
p = ncol(counts)
```
```{r}
head(counts)
head(covar)
```

### Fit PLN model
This creates a `PLNmodels` object
```{r}
library(PLNmodels)
model<-PLN(counts ~ covar$site)
```

### Run EMtree function

```{r output}
library(EMtree)
set.seed(3)
output<-EMtree(model,  maxIter = 10, plot=TRUE)
str(output)
```

### Foster robustness with resampling :


```{r, cache=TRUE}
library(parallel)
resample_output<-ResampleEMtree(counts=counts, covar_matrix = covar$site , S=5, maxIter=10,cond.tol=1e-8, cores=1)
str(resample_output)
```

### Several models with resampling :


```{r, cache=FALSE}
library(parallel)
tested_models=list(1,2,c(1,2))
models_names=c("date","site","date + site")
compare_output<-ComparEMtree(counts, covar_matrix=covar, models=tested_models, m_names=models_names, Pt=0.15,  S=3, maxIter=5,cond.tol=1e-8,cores=1)

str(compare_output)
```
### Graphics
#### From `EMtree` output

Simple network:

```{r, fig.asp=1, fig.width=4.5, fig.fullwidth=FALSE, message=FALSE}
library(ggraph)
library(tidygraph)
library(viridis)

set.seed(200)
edges_prob<- output$edges_prob
edges_prob[edges_prob<2/p]<-0
draw_network(edges_prob,title="Site", pal="dodgerblue3", layout="nicely",curv=0.1)$G

```

#### From `ResampleEMtree` output

```{r, fig.height=4.5, fig.width=4.5}

df<-freq_selec(resample_output$Pmat,Pt=2/p+0.1)

draw_network(df,"Site", layout="nicely")$G

df[which(df<1e-6)]=0
draw_network(df,"Site", layout="nicely")$G
draw_network(df,"Site", layout="nicely")$graph_data

```

#### Facet for plotting several models in one shot

Comparing network by eye is difficult. In particular, choosing the right layout to do so is often troublesome. Here by default, the circle layout is used so that differences in density are easily seen.

```{r, fig.height=3.5, fig.width=10}
compar_graphs(compare_output,alpha=TRUE)$G
```

However, the user can decide another layout. The nodes position is preserved along the networks.

```{r}
compar_graphs(compare_output,alpha=FALSE, layout="nicely", curv=0.1, base_model="site")$G
```


