---
title: "Fatala fishes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fatala fishes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include=FALSE}
library(parallel)
library(ggraph)
library(tidygraph)
library(viridis)
library(tidyr)
library(gridExtra)
library(RColorBrewer)
```

This vignette details how to infer a network, using fishes counts from the Fatala River (Barans95 data from the `ade4` package), and shows how to represent a network with the `EMtree` package.

## Fatala fishes dataset

The data is composed of 33 species abundances measures in 95 samples.

```{r fish_data , message=FALSE}
library(ade4)
library(tibble)
data(baran95)
Y = as.matrix(baran95$fau)
n = nrow(Y)
p = ncol(Y)
head(Y[,1:5])
str(Y)
```

The available covariates are the site and date of the samples. At each date and site, 4 samples were measured, except for kilometer 03 of January 93.
```{r}
X = as_tibble(baran95$plan)
str(X)
table(X$date, X$site)
```

## Network inference

### Basic model

When dealing with abundance data, most modelling approaches switch back to the Gaussian space using different methods. EMtree is an inference procedure which only requires an estimate of a Gaussian covariance matrix, and can be used with any model which either use Gaussian latent variables, Gaussian copulas, or Gaussian data transformations.

The Poisson log-normal model, implemented and estimated in the `PLNmodels` package, is easy to use and can handle both covariates and offsets efficiently. Details on covariates and offsets specification can be found [here](https://pln-team.github.io/PLNmodels/articles/PLN.html). For this example we use PLNmodels, however the user is free to provide a Gaussian correlation estimate as an input.


We create a  `PLNmodels` object:
```{r}
library(PLNmodels)
PLNfit<-PLN(Y ~ X$site)
```



And then run EMtree as follows:

```{r output}
library(EMtree)
EMtreeFit<-EMtree(PLNfit,  maxIter = 20, plot=TRUE)
str(EMtreeFit)
```

The output matrix `edges_prob` gathers the probabilities for each edge to be part of the latent random tree of the model. By construction, this matrix is constrained to sum to $(p-1)$ (number of edges in a spanning tree), and to have an average value of $2/p$.

This package involves a lot of switching between symmetric matrices and the vector of their upper triangular part. The functions `ToVec()` and `ToSym()` help to transition from one representation to the other.

```{r}
vec_of_proba=ToVec(EMtreeFit$edges_prob)
mean(vec_of_proba)-2/p
sum(vec_of_proba) - (p-1)
```

To obtain a network and select edges, we resort to a stability selection strategy.

### Stability selection

The function `ResampleEMtree()` implements a stability selection of EMtree on S sub-samples. This function uses parallel computations with `mclapply()`.  The `Pmat` output gathers all the inferred edges probabilities for each sub-sample ; it is a stack of each `ToVec(EMtreeFit$edges_prob)`.
```{r, cache=TRUE}
ResampEmtreeFit<-ResampleEMtree(counts=Y, covar_matrix = X$site , S=5, maxIter=10,cond.tol=1e-8, cores=1)
str(ResampEmtreeFit)
```

Another way to specify the covariates matrix is to use the model.matrix() function as follows:
```{r}
built_covar_matrix=model.matrix(~X$site+X$date)
```
Then `built_covar_matrix` would be the `covar_matrix` parameter.

The `ResampleEMtree()` function can be used with another model than PLN. To do this, simply provide a function which estimates a Gaussian correlation matrix from counts (using latent variables, copulas or transformations). The function should be passed to `ResampleEMtree()` through the `user_covariance_estimation` parameter. The covariate matrix will also be a parameter to the user's function. An example using a dummy estimation is available in the help of `ResampleEMtree()`.



For the sake of the example we continue with PLN and the "site" covariate only. Edges selection frequencies can be derived from the `Pmat` output with the function `freq_selec()`. This function simply summarizes the number of times an edge has a probability above the threshold `Pt`. A network can then be obtained by thresholding the frequencies, to keep for example edges that are selected in more than $80\%$ of sub-samples:
```{r}
freqs<-freq_selec(ResampEmtreeFit$Pmat,Pt=0.2)
a_first_idea_of_network<-1*(freqs>0.8)
```

The problem of course is that there is no rationale for the threshold on probabilities `Pt` ; the optimal threshold is unknown. A way to select the threshold is to compute the stability of the frequencies for any threshold, and then select the desired stability. The function `StATS()` adapts the well-known penalty selection strategy StARS in the context of threshold selection.
```{r}
stab_selection=StATS(ResampEmtreeFit$Pmat, nlambda=50, stab.thresh=0.9,plot=TRUE)
```

We then keep the frequencies corresponding to the optimal lambda value, which is here at about $exp(-9.5)$. The optimal frequencies are stored in `stab_selection$freqs_opt`, the optimal threshold is `stab_selection$lambda_opt`.

The optimal frequencies still need to be thresholded in order to obtain a network. However, the task is very much simpler to understand now: the higher the frequency, the more stable the edge along the resamples.

## Visualizations

The package `EMtree` has one general plotting function for network visualization: `draw_network()`. It builds from the `ggraph` and `tidygraph` packages, and provides with several interesting functionalities, among which:

  - highlighting of important nodes,
  - shading of edges,
  - using a previously stored layout,
  - defining groups of nodes and edges, 
  - using the igraph output for further graphic purposes.


### Basic usage

For this section we consider three networks from the previous inference: the weighted network with all edges selection frequencies, the network obtained when applying a 0.5 threshold on frequencies, and finally the one obtained with a threshold of 0.95.
```{r}
weighted_net=ToSym(stab_selection$freqs_opt)
net_50=ToSym(1*(stab_selection$freqs_opt>0.5))
net_95=ToSym(1*(stab_selection$freqs_opt>0.95))
```

The function `draw_network()` takes a possibly weighted adjacency matrix as input, and represents a network with edges widths proportional to the input weights. Several layouts are available (among which nicely, kk, fr, circle, stress... See igraph documentation or this [blogpost](https://cran.r-project.org/web/packages/ggraph/vignettes/Layouts.html)). Highlighting of nodes possessing among the highest betweenness centrality measure is done with the parameter `btw_rank`. To disable this functionality, just set `btw_rank=1`.

`shade=TRUE` lowers the opacity of edges not linked to a highlighted node. This allows to better see the direct neighborhood of important nodes.

```{r, fig.height=4, fig.width=8, fig.fullwidth=FALSE, message=FALSE}

g1<-draw_network(weighted_net,title="Weighted", pal_edges="dodgerblue3",layout="stress",
                 shade = TRUE, btw_rank=3)$G
g2<-draw_network(net_50,title="Selected > 50%", pal_edges="dodgerblue3",layout="kk",
                 shade = TRUE, btw_rank=3)$G
g3<-draw_network(net_95,title="Selected > 90%", pal_edges="dodgerblue3",layout="fr",
                 shade = FALSE, btw_rank=3)
grid.arrange(g1, g2, g3$G, ncol=3)


```

### Storing a layout

To store a layout, we need to create it first, using the ggraph package for example. 

```{r}
a_layout = create_layout(net_95, layout="nicely") %>% data.frame()
```

Then the layout is taken as the `stored_layout` parameter of `draw_network()`, as follows:
```{r,fig.height=4, fig.width=7}
args=list( pal_edges="dodgerblue3")
g1<-draw_network(net_50,title="Selected > 50%", pal_edges="dodgerblue3",
                 stored_layout = a_layout, btw_rank=3)$G
g2<-draw_network(net_95,title="Selected > 95%", pal_edges="dodgerblue3",
                 stored_layout = a_layout, btw_rank=3)$G
grid.arrange(g1, g2, ncol=2)
```


### Grouping of nodes and edges

It is possible to define groups of nodes and edges with `node_groups` and `edge_groups` respectively, and color them accordingly using `pal_nodes` and `pal_edges`. For example :

```{r,fig.height=4, fig.width=4}
draw_network(net_95, title="Groups of nodes", pal_edges="gray70", 
             node_groups=sample(3,p, replace=TRUE),
             pal_nodes=brewer.pal(3,"Dark2"),layout="stress")$G
```

```{r,fig.height=4, fig.width=4}
draw_network(net_95, title="Groups of edges", pal_nodes="gray70", 
             edge_groups=ToSym(sample(3,p*(p-1)/2,replace=TRUE)),
             pal_edges=brewer.pal(3,"Dark2"),layout="stress", btw_rank=1)$G
```


### Usage example of the non-graphic output

The `draw_network()` function outputs the graph `G`, but also `graph_data`, which is in igraph format and thus useful if you would like to use your own graphic representation, or another package like `networkD3` for example. 

In the example below, we first create a graph object with nodes labels as the species names. We then extract the edges and nodes information following the tidygraph grammar, and store them in data frames. Edges information are modified so that they are zero-indexed, which is necessary for JavaScript.
Finally we use the `networkD3::forceNetwork()` function to obtain an interactive network, where we colored  nodes according to their degree.
```{r, fig.width=5, fig.height=5}
library(networkD3)
graph<-draw_network(net_95,nodes_label = baran95$species.names)
edges_data=graph$graph_data %>% activate(edges) %>% as_tibble() %>% 
  mutate(from=from-1, to = to-1)  %>% data.frame()
nodes_data=graph$graph_data %>% activate(nodes) %>% data.frame()
forceNetwork(Links = edges_data, Nodes =nodes_data, 
             Source = 'from', Target = 'to', 
             NodeID = 'name', Group = 'deg',opacity = 1,fontSize = 15, legend=TRUE)
```

Another way to do this is to use the `networkD3::igraph_to_networkD3()`, which takes advantage of the igraph nature of the `graph_data` output. It gives the right format and automatically re-index the edges, but requires to manually add the degrees or any other information on node and edges.

```{r, eval=FALSE}
degrees=graph$graph_data %>% activate(nodes) %>% dplyr::select(deg) %>% pull()
graph_d3 <- igraph_to_networkD3(graph$graph_data, group=degrees)
forceNetwork(Links = graph_d3$links, Nodes =graph_d3$nodes, 
             Source = 'source', Target = 'target', 
             NodeID = 'name', Group = 'group',opacity = 1,fontSize = 15, legend=TRUE)
```



