---
title: "Fatala fishes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fatala fishes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include=FALSE}
library(parallel)
library(ggraph)
library(tidygraph)
library(viridis)
```

This is a basic example detailing how to infer a network, using fishes counts from the Fatala River (Barans95 data from the `ade4` package).

## Fatala fishes dataset

The data is composed of 33 species abundances measures in 95 samples.

```{r fish_data , message=FALSE}
library(ade4)
library(tibble)
data(baran95)
Y = as.matrix(baran95$fau)
n = nrow(Y)
p = ncol(Y)
head(Y[,1:5])
str(Y)
```

The available covariates are the site and date of the samples. At each date and site, 4 samples were measured, except for kilometer 03 of January 93.
```{r}
X = as_tibble(baran95$plan)
str(X)
table(X$date, X$site)
```

## Network inference

EMtree infers a network from either a correlation matrix of a multivariate Gaussian, or an object created by PLNmodels from count data.
Therefore here we first create a  `PLNmodels` object:
```{r}
library(PLNmodels)
PLNfit<-PLN(Y ~ X$site)
```

And then run EMtree:

```{r output}
library(EMtree)
EMtreeFit<-EMtree(PLNfit,  maxIter = 20, plot=TRUE)
str(EMtreeFit)
```

To get a network from afit of `EMtree()`, the probabilities stored in `edges_prob` can be thresholded. We propose the $2/p$ threshold as follows:
```{r}
probs<- EMtreeFit$edges_prob
net<-1*(probs>2/p)
```

To improve the robstness, the function `ResampleEMtree()` implements a statibility selection of EMtree on S sub-samples. This function uses parallel computations with `mclapply()`.  The output Pmat gathers all the infered edges probabilities for each sub-sample.
```{r, cache=TRUE}
ResampEmtreeFit<-ResampleEMtree(counts=Y, covar_matrix = X$site , S=5, maxIter=10,cond.tol=1e-8, cores=1)
str(ResampEmtreeFit)
```
Edges selection frequencies can be derived from the `Pmat`output with the function `freq_selec()`. A final network can then be obtained by thresholding the frequencies, to keep for example edges that are selected in more than $80\%$ of sub-samples:
```{r}
freqs<-freq_selec(ResampEmtreeFit$Pmat,Pt=2/p+0.1)
resampNet<-1*(freqs>0.8)
```

## Infer networks under several models:

The aim of function `ComparEMtree()` is to run network inference with different covariates specifications. It uses `ResampleEMtree()` and adjust the different models specified in `model_names` as follows: 
```{r, cache=FALSE}
tested_models=list(1,2,c(1,2))
models_names=c("date","site","date + site")
compare_models<-ComparEMtree(Y, X, models=tested_models, m_names=models_names, Pt=2/p+0.1,  S=3, maxIter=5,cond.tol=1e-8,cores=1)

str(compare_models)
```

## Visuals

Package Emtree provides with plotting functions for network visualizations. They build from the `ggraph` and `'tidygraph` packages.

### Simple networks
The function `draw_network()` takes a weighted matrix an input, and represent a network with edges widths proportional to the input weights. Several layouts are available (see the `ggraph` documentation). Highlighting of nodes possessing among the highest betweenness centrality measure is done with the parameter `btw_rank`.

 

```{r, fig.asp=1, fig.width=4.5, fig.fullwidth=FALSE, message=FALSE}
probs[probs<2/p]=0 
draw_network(probs,title="Site", pal_edges="dodgerblue3", layout="nicely",btw_rank=3)$G
```

From the selected edges frequencies:
```{r, fig.height=4.5, fig.width=4.5}
draw_network(resampNet,"Site", layout="nicely")$G 

```

### Facets of several networks

The function`compare_graphs()` draws a facet plot of the output networks from `ComparEMtree()`.  Comparing network by eye is difficult, on particular choosing the right layout to do so is often troublesome. Here by default, the circle layout is used so that differences in density and sensitive nodes are easily seen.

```{r,  fig.height=3.5, fig.width=8}
compare_graphs(compare_models,shade=TRUE)$G
```

However, if another layout is preferred, the nodes position is preserved along the facet and defined by choosing the `base_model`.

```{r,  fig.height=3.5,fig.width=8}
compare_graphs(compare_models,shade=FALSE, layout="nicely", curv=0.1, base_model="site")$G
```


